Evolution of \\ ice sheet geometry \\ using Stokes dynamics

### 2021-06-02 talk/talk.tex
Progress slides.

@@

\begin{itemize}
\item \alert{a nonlinear complementarity problem (NCP) coupled to the Glen-Stokes model for ice velocity $\bu$:}

\vspace{-2mm}
\begin{align*}
s - b &\ge 0 && \text{on $\Omega$} \\
- \bu|_s \cdot \bn_s - a &\ge 0 && \text{''} \\
(s - b) (- \bu|_s \cdot \bn_s - a) &= 0 && \text{''} \\
- \nabla \cdot \left(2 \nu_\eps\, D\bu\right) + \nabla p - \rhoi \mathbf{g} &= \bzero && \text{on $\Lambda_s$} \\
\nabla \cdot \bu &= 0 && \text{''} \\
\left(2 \nu_\eps D\bu - pI\right) \bn &= \bzero && \text{on $\partial \Lambda_s \setminus \Gamma_0$} \\
\bu &= \bzero && \text{on $\Gamma_0$ (ice base)}
\end{align*}

    \begin{itemize}
    \item Glen-law effective viscosity with $\text{p}=(1/\text{n})+1=4/3$:
      $$\nu_\eps = \frac{1}{2} B_n \left(|D\bu|^2 + \eps\, D_0^2\right)^{(\pp-2)/2}$$
    \item surface kinematical equation holds \emph{on the ice}: \quad $\bu|_s \cdot \bn_s + a=0$
    \end{itemize}
\end{itemize}
\end{frame}

@@

\begin{itemize}
\item almost no one is solving the SIGP directly, or implicit time-stepping
    \begin{itemize}
    \item except Wirbel \& Jarosch \cite{WirbelJarosch2020} \dots nonscalable, ``fake ice''
    \item optimal fixed-geometry Stokes solver (Isaac and others \cite{IsaacStadlerGhattas2015})
    \end{itemize}
%\item \emph{warning}: barrier to exact numerical mass conservation \cite{Bueler2021conservation}
\item what are people doing?
    \begin{itemize}
    \item \alert{explicit time-stepping with $s \ge b$ enforced by truncation} (below)
    \end{itemize}
\end{itemize}

\medskip
\begin{columns}
    \column{0.8\textwidth}
        \hfill \includegraphics[width=0.8\textwidth]{figs/seguinot.png}
    \column{0.2\textwidth}
        \scriptsize \emph{show clip from Julien Seguinot using PISM}
\end{columns}
\end{frame}

@@

@@

\begin{frame}{References}
	\bibliography{../paper/msg.bib}
	\bibliographystyle{plain}
\end{frame}

>>>>

\usepackage{tikz}
\usetikzlibrary{decorations.pathreplacing,calc}

\newcommand{\tikzmark}[1]{\tikz[overlay,remember picture] \node (#1) {};}

@@

\begin{itemize}
\item \alert{a nonlinear complementarity problem (NCP) coupled to the Glen-Stokes model for $\bu,p$:}

\vspace{-5mm}

\begin{align*}
s - b &\ge 0 && \text{on $\Omega$} & \tikzmark{first} \\
- \bu|_s \cdot \bn_s - a &\ge 0 && \text{''} \\
(s - b) (- \bu|_s \cdot \bn_s - a) &= 0 && \text{''} & \tikzmark{second} \\
- \nabla \cdot \left(2 \nu_\eps\, D\bu\right) + \nabla p - \rhoi \mathbf{g} &= \bzero && \text{on $\Lambda_s$} \\
\nabla \cdot \bu &= 0 && \text{''} \\
\bu &= \bzero && \text{on $\Gamma_0$ \quad (ice base)} \\
\left(2 \nu_\eps D\bu - pI\right) \bn &= \bzero && \text{on $\partial \Lambda_s \setminus \Gamma_0$}
\end{align*}

\begin{tikzpicture}[overlay, remember picture]
\draw[decoration={brace,amplitude=0.4em},decorate,ultra thick] (first.north east) node[right=8mm,below=3mm] {NCP} -- (second.east);
\end{tikzpicture}

\vspace{-2mm}

    \begin{itemize}
    \item $\bn_s = \left<-s_x,-s_y,1\right>$  \quad (normal to ice surface; upward)
    \item surface kinematical equation holds \emph{on the ice}: \quad $\bu|_s \cdot \bn_s + a=0$
    \item Glen-law effective viscosity with $\text{p}=(1/\text{n})+1=4/3$:
      $$\nu_\eps = \frac{1}{2} B_n \left(|D\bu|^2 + \eps\, D_0^2\right)^{(\pp-2)/2}$$
    \end{itemize}
\end{itemize}
\end{frame}

@@

\begin{itemize}
\item almost no one is solving the SIGP directly, or doing implicit time-stepping
    \begin{itemize}
    \item except Wirbel \& Jarosch \cite{WirbelJarosch2020} \dots nonscalable, w.~``fake ice''
    \end{itemize}
\item what are people doing?
    \begin{itemize}
    \item \alert{explicit time-stepping with $s \ge b$ enforced by truncation} (below)
    \end{itemize}
\end{itemize}

\medskip
\begin{columns}
    \column{0.8\textwidth}
        \hfill \includegraphics[width=0.8\textwidth]{figs/seguinot.png}
    \column{0.2\textwidth}
        \scriptsize \emph{show clip of PISM run by Julien Seguinot}
\end{columns}
\end{frame}

\begin{frame}{goals}

\begin{itemize}
\item FIXME
\end{itemize}
\end{frame}

\begin{frame}{ice dynamics operator}

\begin{itemize}
\item FIXME
\item optimal fixed-geometry Stokes solver exists (Isaac et al \cite{IsaacStadlerGhattas2015})
\end{itemize}
\end{frame}

@@

\begin{frame}{extra}
 NCP is a barrier to exact numerical mass conservation \cite{Bueler2021conservation}
\end{frame}

@@

\begin{frame}{References}

%\bibliography{../paper/msg.bib}
%\bibliographystyle{plain}

\begin{thebibliography}{1}
\bibitem{Bueler2021conservation}
E.~Bueler. \textbf{Conservation laws for free-boundary fluid layers.} {\em SIAM J. Appl. Math.}, to appear.

\bibitem{BuelerMitchell2022}
E.~Bueler and L.~Mitchell. \textbf{Multilevel computation of glacier geometry from Stokes dynamics.} in preparation

\bibitem{IsaacStadlerGhattas2015}
T.~Isaac, G.~Stadler, and O.~Ghattas.  \textbf{Solution of nonlinear Stokes equations discretized by high-order finite elements on nonconforming and anisotropic meshes, with application to ice sheet dynamics.} {\em SIAM J. Sci. Comput.}, 37(6):B804--B833, 2015.

\bibitem{WirbelJarosch2020}
A.~Wirbel and A.~H. Jarosch. \textbf{Inequality-constrained free-surface evolution in a full {S}tokes ice flow model (\textit{evolve\_glacier v1.1}).} {\em Geoscientific Model Development}, 13(12):6425--6445, 2020.
\end{thebibliography}

\end{frame}
